---
title: "A Report Lake Ontario's Microbes"
author: "Hilal Samut"
date: "2025-02-19"
output:
  html_document:
    df_print: paged
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r packages}
# Load libraries/packages
library(tidyverse)
```


# Load in the Lake Ontario Data

```{r load-data}
# Load in lake ontario microbial community data
sample_and_taxon <-
  read_csv("data/sample_and_taxon.csv")

# Inspect the data
glimpse(sample_and_taxon)
```


# Microbial Abundance Versus Temperature

```{r cells_vs_temp}
# temp on the X
# cell abundance on th Y
# Colored by environmental Group
# make it look nice e.g. labels, models etc
ggplot(data = sample_and_taxon,
       aes(x = temperature, y = cells_per_ml/10^6)) +
  geom_point(aes(color = env_group)) + 
  labs(x = "Temperature (C)", 
       y = "Cell Abundance (cfu/ml)") +
  scale_color_manual(values = c("cornflowerblue",
                                "goldenrod1",
                                "firebrick")) +
  geom_smooth(method = "lm", color = "black") +
  theme_classic() +
  theme(legend.position = "bottom", 
        legend.title = element_blank())
``` 

The above plot shows that:

- There's a positive relationship between temperature and cell abundances.
- For example, deep samples are the coldest and have the fewest cells. 

```{r data-summary, include = FALSE}
n_samples <-
  sample_and_taxon %>%
  #Only pick out sample id column
  select(sample_id) %>%
  # Number of distinct names in the column 
  n_distinct()

# Calculate the minimum and maximum temps
# min
minTemp <- 
  sample_and_taxon %>% 
  summarize(round(min(temperature), digits = 1)) %>%
  pull()

# max
maxTemp <- 
  sample_and_taxon %>% 
  summarize(round(max(temperature), digits = 1)) %>%
  pull()
```

The total number of samples is `r n_samples`. For this set of samples, temperature ranged from a minimum of `r minTemp` to a `r maxTemp` celcius. 
